---
title: "AMR carriage in urban wild small mammals"
date: "2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 1. Packages

```{r packages}
library(tidyverse)
library(janitor)
library(rstatix)
library(broom)
library(geepack)
library(MuMIn)
library(brglm2)
library(gtsummary)
library(flextable)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(pheatmap)
library(grid)
library(gtable)
library(gridExtra)
library(svglite)
library(AMR)
```

# 2. Data

```{r data}
Data_indiv <- read.csv("Data_indiv.csv", sep = ";")
Data_RAM   <- read.csv("Data_antibiograms.csv", sep = ";")
```

# 3. Antibiogram processing

```{r sir-conversion}
sir_cols <- names(Data_RAM)[grepl("_SIR$", names(Data_RAM))]
Data_RAM <- Data_RAM %>%
  mutate(across(all_of(sir_cols), as.sir))
```

```{r sir-to-num}
sir_to_num <- function(x) {
  case_when(x == "S" ~ 0, x == "I" ~ 1, x == "R" ~ 2, TRUE ~ NA_real_)
}
```

# 4. Resistance profile heatmaps

## 4.1 ESBL Enterobacteriaceae (E. coli and K. pneumoniae)

```{r heatmap-esbl}
blse <- Data_RAM %>%
  filter(Culture_medium == "BLSE",
         Bacterial_species %in% c("Klebsiella pneumoniae", "Escherichia coli")) %>%
  select(Specimen_id, Bacterial_species,
         Amox25_SIR, AUG30_SIR, PRL100_SIR, PTZ36_SIR,
         CFX30_SIR, EFT30_SIR, CEQ30_SIR,
         IMI10_SIR, ETP_SIR,
         S10_SIR, K30_SIR, GM15_SIR,
         FLU3_SIR, ENF5_SIR, MAR5_SIR,
         TS25_SIR, FFC30_SIR, CO50_SIR) %>%
  mutate(n_R = rowSums(across(ends_with("_SIR"), ~ . == "R"), na.rm = TRUE)) %>%
  group_by(Specimen_id) %>%
  slice_max(n_R, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-n_R)

mat_data <- blse %>% mutate(across(ends_with("_SIR"), sir_to_num))

row_labels <- paste0(
  gsub("BLSE ", "", mat_data$Specimen_id), "_",
  ifelse(mat_data$Bacterial_species == "Klebsiella pneumoniae", "Kp", "Ec")
)

mat <- mat_data %>% select(ends_with("_SIR")) %>% as.matrix()
rownames(mat) <- row_labels
colnames(mat) <- c("AMX", "AUG", "PIP", "TZP",
                   "CFX", "EFT", "CEQ",
                   "IMI", "ETP",
                   "STR", "KAN", "GM",
                   "FLU", "ENF", "MAR",
                   "SXT", "FFC", "COL")

annotation_row <- data.frame(
  Species = ifelse(grepl("Kp", row_labels), "K. pneumoniae", "E. coli"),
  row.names = row_labels
)

annotation_col <- data.frame(
  `Antibiotic class` = c(
    "Penicillin", "Penicillin + BLI", "Ureidopenicillin", "Ureidopenicillin + BLI",
    "Cephalosporin", "Cephalosporin", "Cephalosporin",
    "Carbapenem", "Carbapenem",
    "Aminoglycoside", "Aminoglycoside", "Aminoglycoside",
    "Fluoroquinolone", "Fluoroquinolone", "Fluoroquinolone",
    "Sulfonamide", "Phenicol", "Polymyxin"),
  check.names = FALSE, row.names = colnames(mat)
)

ann_colors <- list(
  Species = c("K. pneumoniae" = "#2166AC", "E. coli" = "#F527F2"),
  `Antibiotic class` = c(
    "Penicillin" = "#FDB462", "Penicillin + BLI" = "#FEE391",
    "Ureidopenicillin" = "#8DD3C7", "Ureidopenicillin + BLI" = "#80B1D3",
    "Cephalosporin" = "#FB8072", "Carbapenem" = "#BEBADA",
    "Aminoglycoside" = "#B3DE69", "Fluoroquinolone" = "#FCCDE5",
    "Sulfonamide" = "#D9D9D9", "Phenicol" = "#BC80BD", "Polymyxin" = "#A6CEE3"
  )
)

ordre_lignes <- c(grep("Ec", row_labels, value = TRUE),
                  grep("Kp", row_labels, value = TRUE))
mat            <- mat[ordre_lignes, ]
annotation_row <- annotation_row[ordre_lignes, , drop = FALSE]

couleurs_sir <- c("#2CA02C", "#FFD700", "#D62728")

p_blse <- pheatmap(
  mat, color = couleurs_sir, breaks = c(-0.5, 0.5, 1.5, 2.5),
  annotation_row = annotation_row, annotation_col = annotation_col,
  annotation_colors = ann_colors,
  cluster_rows = FALSE, cluster_cols = FALSE,
  gaps_col = c(1, 2, 3, 4, 7, 9, 12, 15, 16, 17),
  fontsize = 10, fontsize_row = 9, fontsize_col = 9, angle_col = 45,
  legend_breaks = c(0, 1, 2), legend_labels = c("S", "I", "R"),
  na_col = "#F0F0F0", main = "Resistance profiles — ESBL Enterobacteriaceae",
  silent = TRUE
)

png("heatmap_ESBL.png", width = 14, height = 7, units = "in", res = 300)
grid::grid.newpage(); grid::grid.draw(p_blse$gtable); dev.off()
```

## 4.2 Staphylococcus aureus (SAID and MRSA media)

```{r heatmap-saureus}
sa <- Data_RAM %>%
  filter(Culture_medium %in% c("SAID", "SARM"),
         Bacterial_species == "Staphylococcus aureus") %>%
  select(Specimen_id, Culture_medium,
         Peni41_SIR, AUG30_SIR, FOX30_SIR, CEF30_SIR,
         S10_SIR, GM15_SIR, K30_SIR,
         MAR5_SIR, E15_SIR, L15_SIR,
         TS25_SIR, T30_SIR, RP30_SIR) %>%
  mutate(n_R = rowSums(across(ends_with("_SIR"), ~ . == "R"), na.rm = TRUE)) %>%
  group_by(Specimen_id, Culture_medium) %>%
  slice_max(n_R, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-n_R)

mat_data_sa <- sa %>% mutate(across(ends_with("_SIR"), sir_to_num))

row_labels_sa <- paste0(
  gsub("SAID |SARM ", "", mat_data_sa$Specimen_id), "_",
  ifelse(mat_data_sa$Culture_medium == "SAID", "SAID", "MRSA")
)

mat_sa <- mat_data_sa %>% select(ends_with("_SIR")) %>% as.matrix()
rownames(mat_sa) <- row_labels_sa
colnames(mat_sa) <- c("PEN", "AUG", "FOX", "CFX",
                      "STR", "GM", "KAN", "MAR",
                      "ERY", "CLI", "SXT", "TET", "RIF")

annotation_row_sa <- data.frame(
  Medium = ifelse(grepl("SAID$", row_labels_sa), "SAID", "MRSA"),
  row.names = row_labels_sa
)

annotation_col_sa <- data.frame(
  `Antibiotic class` = c(
    "Penicillin", "Penicillin + BLI",
    "Cephalosporin", "Cephalosporin",
    "Aminoglycoside", "Aminoglycoside", "Aminoglycoside",
    "Fluoroquinolone", "Macrolide", "Lincosamide",
    "Sulfonamide", "Tetracycline", "Rifamycin"),
  check.names = FALSE, row.names = colnames(mat_sa)
)

ann_colors_sa <- list(
  Medium = c("SAID" = "#F132EB", "MRSA" = "#2166AC"),
  `Antibiotic class` = c(
    "Penicillin" = "#FDB462", "Penicillin + BLI" = "#FEE391",
    "Cephalosporin" = "#FB8072", "Aminoglycoside" = "#B3DE69",
    "Fluoroquinolone" = "#FCCDE5", "Macrolide" = "#DFC27D",
    "Lincosamide" = "#BC80BD", "Sulfonamide" = "#D9D9D9",
    "Tetracycline" = "#CCEBC5", "Rifamycin" = "#80B1D3"
  )
)

ordre_lignes_sa <- c(grep("SAID$", row_labels_sa, value = TRUE),
                     grep("MRSA$", row_labels_sa, value = TRUE))
mat_sa            <- mat_sa[ordre_lignes_sa, ]
annotation_row_sa <- annotation_row_sa[ordre_lignes_sa, , drop = FALSE]

p_sa <- pheatmap(
  mat_sa, color = couleurs_sir, breaks = c(-0.5, 0.5, 1.5, 2.5),
  annotation_row = annotation_row_sa, annotation_col = annotation_col_sa,
  annotation_colors = ann_colors_sa,
  cluster_rows = FALSE, cluster_cols = FALSE,
  gaps_col = c(1, 2, 4, 7, 8, 10, 11, 12),
  fontsize = 10, fontsize_row = 9, fontsize_col = 9, angle_col = 45,
  legend_breaks = c(0, 1, 2), legend_labels = c("S", "I", "R"),
  na_col = "#F0F0F0", main = "Resistance profiles — Staphylococcus aureus",
  silent = TRUE
)

png("heatmap_Saureus.png", width = 9, height = 11, units = "in", res = 300)
grid::grid.newpage(); grid::grid.draw(p_sa$gtable); dev.off()
```

## 4.3 Acinetobacter spp. (A. baumannii and A. calcoaceticus)

```{r heatmap-acinetobacter}
acba <- Data_RAM %>%
  filter(Culture_medium == "ACBA",
         Bacterial_species %in% c("Acinetobacter baumannii",
                                  "Acinetobacter calcoaceticus")) %>%
  select(Specimen_id, Bacterial_species,
         AUG30_SIR, CEQ30_SIR, CPM30_SIR, CAZ30_SIR,
         PRL100_SIR, PTZ36_SIR, IMI10_SIR,
         GM10_SIR, AK30_SIR, CO50_SIR,
         TS25_SIR, T30_SIR, CIP5_SIR, LEV5_SIR) %>%
  mutate(n_R = rowSums(across(ends_with("_SIR"), ~ . == "R"), na.rm = TRUE)) %>%
  group_by(Specimen_id) %>%
  slice_max(n_R, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-n_R)

mat_data_acba <- acba %>% mutate(across(ends_with("_SIR"), sir_to_num))

row_labels_acba <- paste0(
  gsub("ACBA ", "", mat_data_acba$Specimen_id), "_",
  ifelse(mat_data_acba$Bacterial_species == "Acinetobacter baumannii", "Ab", "Ac")
)

mat_acba <- mat_data_acba %>% select(ends_with("_SIR")) %>% as.matrix()
rownames(mat_acba) <- row_labels_acba
colnames(mat_acba) <- c("AMC", "CEQ", "CPM", "CAZ",
                        "PRL", "TZP", "IMI",
                        "GM", "AMK", "COL",
                        "SXT", "TET", "CIP", "LEV")

annotation_row_acba <- data.frame(
  Species = ifelse(grepl("Ab$", row_labels_acba),
                   "A. baumannii", "A. calcoaceticus"),
  row.names = row_labels_acba
)

annotation_col_acba <- data.frame(
  `Antibiotic class` = c(
    "Penicillin + BLI",
    "Cephalosporin", "Cephalosporin", "Cephalosporin",
    "Ureidopenicillin", "Ureidopenicillin + BLI",
    "Carbapenem", "Aminoglycoside", "Aminoglycoside",
    "Polymyxin", "Sulfonamide", "Tetracycline",
    "Fluoroquinolone", "Fluoroquinolone"),
  check.names = FALSE, row.names = colnames(mat_acba)
)

ann_colors_acba <- list(
  Species = c("A. baumannii" = "#3C27F5", "A. calcoaceticus" = "#F527F5"),
  `Antibiotic class` = c(
    "Penicillin + BLI" = "#FDB462",
    "Cephalosporin" = "#FB8072",
    "Ureidopenicillin" = "#8DD3C7",
    "Ureidopenicillin + BLI" = "#80B1D3",
    "Carbapenem" = "#BEBADA", "Aminoglycoside" = "#B3DE69",
    "Polymyxin" = "#CCEBC5", "Sulfonamide" = "#D9D9D9",
    "Tetracycline" = "#BC80BD", "Fluoroquinolone" = "#FCCDE5"
  )
)

ordre_lignes_acba <- c(grep("Ab$", row_labels_acba, value = TRUE),
                       grep("Ac$", row_labels_acba, value = TRUE))
mat_acba            <- mat_acba[ordre_lignes_acba, ]
annotation_row_acba <- annotation_row_acba[ordre_lignes_acba, , drop = FALSE]

p_acba <- pheatmap(
  mat_acba, color = couleurs_sir, breaks = c(-0.5, 0.5, 1.5, 2.5),
  annotation_row = annotation_row_acba, annotation_col = annotation_col_acba,
  annotation_colors = ann_colors_acba,
  cluster_rows = FALSE, cluster_cols = FALSE,
  gaps_col = c(1, 4, 6, 7, 9, 10, 11, 12),
  fontsize = 10, fontsize_row = 9, fontsize_col = 9, angle_col = 45,
  legend_breaks = c(0, 1, 2), legend_labels = c("S", "I", "R"),
  na_col = "#F0F0F0", main = "Resistance profiles — Acinetobacter spp.",
  silent = TRUE
)

png("heatmap_Acinetobacter.png", width = 10, height = 5, units = "in", res = 300)
grid::grid.newpage(); grid::grid.draw(p_acba$gtable); dev.off()
```

# 5. ESBL carriage prevalence by site type

## 5.1 Observed prevalence with 95% Clopper-Pearson CI

```{r figure-prevalence}
prev_data <- Data_indiv %>%
  filter(Species == "Rattus norvegicus") %>%
  group_by(Loc_type) %>%
  summarise(n_pos = sum(Entero.BLSE, na.rm = TRUE), n_total = n(), .groups = "drop") %>%
  rowwise() %>%
  mutate(
    prev    = n_pos / n_total * 100,
    ic_low  = binom.test(n_pos, n_total)$conf.int[1] * 100,
    ic_high = binom.test(n_pos, n_total)$conf.int[2] * 100,
    label   = paste0(n_pos, "/", n_total)
  ) %>%
  ungroup()

prev_data$Loc_type <- factor(prev_data$Loc_type,
  levels = c("Sewage", "City", "Park", "Zoo"),
  labels = c("Sewer", "Residential", "Green space", "Zoo"))

couleurs <- c("Sewer" = "#08519C", "Residential" = "#6BAED6",
              "Green space" = "#1B9E77", "Zoo" = "#66A61E")

p_prev <- ggplot(prev_data, aes(x = Loc_type, y = prev, fill = Loc_type)) +
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = ic_low, ymax = ic_high),
                width = 0.08, linewidth = 0.9, color = "#333333") +
  geom_text(aes(y = ic_high + 3, label = label), fontface = "bold", size = 4.5) +
  scale_fill_manual(values = couleurs) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) +
  labs(
    title    = expression(paste("ESBL Enterobacteriaceae carriage prevalence in ",
                                italic("Rattus norvegicus"))),
    subtitle = "95% Clopper-Pearson CI",
    x = NULL, y = "Observed prevalence (%)"
  ) +
  theme_classic(base_size = 13) +
  theme(legend.position = "none",
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10, color = "grey40", face = "italic"),
        plot.title.position = "plot")

ggsave("esbl_prevalence.png", p_prev, width = 18, height = 12, units = "cm", dpi = 300)
```

## 5.2 Fisher's exact test

```{r fisher-esbl}
table_BLSE <- table(Data_indiv$Entero.BLSE, Data_indiv$Loc_type)
fisher.test(table_BLSE)
chisq.test(table_BLSE)$stdres

table_BLSE_CP <- table(
  filter(Data_indiv, Loc_type %in% c("City", "Park"))$Entero.BLSE,
  filter(Data_indiv, Loc_type %in% c("City", "Park"))$Loc_type
)
fisher.test(table_BLSE_CP)
```

## 5.3 Site-level carriage

```{r portage-site}
portage_site <- Data_indiv %>%
  filter(Loc_type %in% c("City", "Sewage", "Park", "Zoo")) %>%
  group_by(Loc_number, Loc_type) %>%
  summarise(
    n_total  = n(),
    n_ESBL   = sum(Entero.BLSE, na.rm = TRUE),
    site_pos = as.integer(n_ESBL > 0),
    .groups  = "drop"
  )

portage_site %>%
  group_by(Loc_type) %>%
  summarise(sites_pos = sum(site_pos), sites_tot = n(),
            pct = sites_pos / sites_tot * 100)
```

# 6. Human Footprint Index (HFP-100)

## 6.1 Distribution by site type

```{r hfp-description}
hfp_site <- Data_indiv %>% distinct(Loc_number, Loc_type, HFP.100)

hfp_site %>%
  group_by(Loc_type) %>%
  summarise(n = n(), median = median(HFP.100, na.rm = TRUE),
            mean = mean(HFP.100, na.rm = TRUE), sd = sd(HFP.100, na.rm = TRUE))

hfp_site$Loc_type <- factor(hfp_site$Loc_type,
  levels = c("Sewage", "City", "Park", "Zoo"),
  labels = c("Sewer", "Residential", "Green space", "Zoo"))

kruskal_res <- kruskal.test(HFP.100 ~ Loc_type, data = hfp_site)
print(kruskal_res)

dunn_res <- hfp_site %>% dunn_test(HFP.100 ~ Loc_type, p.adjust.method = "bonferroni")
print(dunn_res)
```

## 6.2 Boxplot

```{r figure-hfp}
dunn_sig <- dunn_res %>%
  filter(p.adj < 0.05) %>%
  mutate(y.position = seq(
    from = max(hfp_site$HFP.100, na.rm = TRUE) * 1.05,
    by   = max(hfp_site$HFP.100, na.rm = TRUE) * 0.08,
    length.out = n()
  ))

p_hfp <- ggplot(hfp_site, aes(x = Loc_type, y = HFP.100, fill = Loc_type)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.8) +
  geom_jitter(width = 0.12, size = 2.5, color = "grey30", alpha = 0.7) +
  scale_fill_manual(values = c("Sewer" = "#08519C", "Residential" = "#6BAED6",
                               "Green space" = "#1B9E77", "Zoo" = "#66A61E"),
                    guide = "none") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.15))) +
  labs(title    = "Human Footprint Index (HFP-100) by site category",
       subtitle = paste0("Kruskal-Wallis: p = ", signif(kruskal_res$p.value, 3)),
       x = NULL, y = "HFP-100") +
  theme_classic(base_size = 13) +
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10, color = "grey40", face = "italic"),
        plot.title.position = "plot")

if (nrow(dunn_sig) > 0)
  p_hfp <- p_hfp + stat_pvalue_manual(dunn_sig, label = "p.adj.signif")

ggsave("HFP100_boxplot.png", p_hfp, width = 18, height = 12, units = "cm", dpi = 300)
```

# 7. Statistical models

## 7.1 Data preparation

```{r data-prep-models}
names(Data_indiv) <- trimws(names(Data_indiv))
names(Data_indiv)[names(Data_indiv) == "HFP.100"]     <- "HFP"
names(Data_indiv)[names(Data_indiv) == "Entero.BLSE"] <- "Entero_BLSE"
names(Data_indiv)[names(Data_indiv) == "Staph.macro"] <- "Staph_macro"

Data_indiv$site        <- as.factor(Data_indiv$Loc_number)
Data_indiv$HFP         <- as.numeric(Data_indiv$HFP)
Data_indiv$Entero_BLSE <- as.integer(Data_indiv$Entero_BLSE)
Data_indiv$Staph_macro <- as.integer(Data_indiv$Staph_macro)
Data_indiv$HFP_scaled  <- scale(Data_indiv$HFP)

Data_rats <- Data_indiv %>%
  filter(Species == "Rattus norvegicus") %>%
  arrange(site)
```

## 7.2 GEE — ESBL carriage ~ HFP-100

```{r gee-esbl}
gee_entero <- geeglm(
  Entero_BLSE ~ HFP_scaled,
  data = Data_rats, family = binomial(link = "logit"),
  id = site, corstr = "exchangeable"
)
summary(gee_entero)
exp(coef(gee_entero)["HFP_scaled"])
```

## 7.3 GEE — macrolide-resistant S. aureus carriage ~ HFP-100

```{r gee-staph}
gee_staph <- geeglm(
  Staph_macro ~ HFP_scaled,
  data = Data_rats, family = binomial(link = "logit"),
  id = site, corstr = "exchangeable"
)
summary(gee_staph)
exp(coef(gee_staph)["HFP_scaled"])
```

## 7.4 Correlation structure selection (QIC)

```{r gee-qic}
gee_indep <- geeglm(Entero_BLSE ~ HFP_scaled, data = Data_rats,
                    family = binomial(link = "logit"), id = site, corstr = "independence")
gee_ar1   <- geeglm(Entero_BLSE ~ HFP_scaled, data = Data_rats,
                    family = binomial(link = "logit"), id = site, corstr = "ar1")
QIC(gee_indep); QIC(gee_entero); QIC(gee_ar1)
```

## 7.5 GEE diagnostics

```{r gee-diagnostics}
table(Data_rats$site)
length(unique(Data_rats$site))

res_pearson <- residuals(gee_entero, type = "pearson")
plot(fitted(gee_entero), res_pearson,
     xlab = "Fitted values", ylab = "Pearson residuals",
     main = "Pearson residuals — GEE ESBL")
abline(h = 0, col = "red", lty = 2)

summary(gee_entero)$coefficients[, c("Std.err", "Naive S.E.")]
```

## 7.6 GEE — individual-level factors (weight, sex)

```{r gee-individual}
Data_rats_complete <- Data_rats %>%
  filter(!is.na(Sex), !is.na(Weight), Sex %in% c("M", "F")) %>%
  mutate(Weight_scaled = scale(Weight), Sex = as.factor(Sex)) %>%
  arrange(site)

for (outcome in c("Entero_BLSE", "Staph_macro")) {
  for (predictor in c("Weight_scaled", "Sex")) {
    f <- as.formula(paste(outcome, "~", predictor))
    m <- geeglm(f, data = Data_rats_complete,
                family = binomial(link = "logit"),
                id = site, corstr = "exchangeable")
    cat("\n---", outcome, "~", predictor, "---\n")
    print(summary(m))
  }
}
```

## 7.7 Firth logistic regression — sector effect at site level

```{r glm-firth}
data_site <- Data_indiv %>%
  filter(Species == "Rattus norvegicus") %>%
  mutate(Sector = case_when(
    Loc_type %in% c("City", "Sewage") ~ "Urban",
    Loc_type %in% c("Park", "Zoo")    ~ "Natural"
  )) %>%
  group_by(Sector, Loc_type, Loc_number, Loc_name) %>%
  summarise(n_rats = n(), n_pos = sum(Entero.BLSE, na.rm = TRUE),
            carriage = as.integer(n_pos > 0), .groups = "drop")

model_firth <- glm(
  carriage ~ Sector + offset(log(n_rats)),
  data = data_site, family = binomial(link = "logit"),
  method = "brglmFit"
)
summary(model_firth)
print(round(cbind(OR = exp(coef(model_firth)), exp(confint(model_firth))), 3))

model_null <- glm(carriage ~ offset(log(n_rats)), data = data_site,
                  family = binomial, method = "brglmFit")
anova(model_null, model_firth, test = "LRT")
```

# 8. Capture effort by site

```{r capture-effort}
counts_per_site <- Data_indiv %>%
  filter(Species == "Rattus norvegicus") %>%
  group_by(Loc_type, Loc_number) %>%
  summarise(n_rats = n(), .groups = "drop")

kw <- kruskal.test(n_rats ~ Loc_type, data = counts_per_site)
print(kw)

kw_label <- paste0("Kruskal-Wallis\nH = ", round(kw$statistic, 2),
                   ", df = ", kw$parameter, ", p = ", signif(kw$p.value, 3))

p_capture <- ggplot(counts_per_site,
                    aes(x = Loc_type, y = n_rats, fill = Loc_type)) +
  geom_boxplot(outlier.shape = 21, outlier.fill = "white", width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2) +
  annotate("text", x = Inf, y = Inf, label = kw_label,
           hjust = 1.05, vjust = 1.5, size = 3.5, color = "grey30") +
  scale_fill_brewer(palette = "Set2") +
  labs(title   = expression(italic("Rattus norvegicus") ~ "— captures per site"),
       x       = "Site type", y = "Number of individuals captured",
       caption = "Kruskal-Wallis test; dots = individual sites") +
  theme_classic(base_size = 13) +
  theme(legend.position = "none")

ggsave("capture_effort.png", p_capture, width = 18, height = 12, units = "cm", dpi = 300)
```
